Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-23 paper/paper.Rmd

We next turn to an assessment of the replication $p$-value to determine whether the key statistic found a significant effect at the traditional $p=.05$ level. Only `r sum(d$rep_p_value_05)` of the studies yielded a significant replication $p$-value. The well-documented "dance of the p-values" [@cumming2013] suggests that the inference drawn from a single replication *p*-value may not be  informative, however, and other metrics such as effect sizes and Bayes factors may provide a more nuanced perspective. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpretating confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

### Effect size and Bayes Factor analysis

We first compared effect sizes from the original and replication studies (Figure 2A). Effect sizes were highly correlated between the two sets of studies 
(r = `r signif(cor.test(d$original_ES_d,d$rep_es)$estimate, digits =2)`, p < .0001), but they were generally smaller in the replications 
(`r signif(coef(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)`\% of the original [`r signif(confint(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)` - `r signif(confint(lm(d$rep_es~d$original_ES_d-1))[2], digits = 2)`]). This estimate should be viewed with caution, however, due to its dependence on one very large effect size estimate ($d > 15$) from a single study.

Next, following @etz2016, we compared Bayes Factors between the original and replication (Figure 2B). The Bayes Factor is an alternative to classical null hypothesis significance testing; it quantifies the evidence in favor of a hypothesis as the ratio of the likelihood of the data under that hypothesis relative to another [@jeffreys1961]. Its proponents point to many advantages over classical $p$-values. For example, it does not privilege the null hypothesis and can provide evidence either *for* or *against* it [see also @scheibehenne2016, @wagenmakers2016]. While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

>>>>

We next turn to an assessment of the replication $p$-value to determine whether the key statistic found a significant effect at the traditional $p=.05$ level. Only `r sum(d$rep_p_value_05)` of the studies yielded a significant replication $p$-value. The well-documented "dance of the p-values" [@cumming2013] suggests that the inference drawn from a single replication p-value may not be  informative, however, and other metrics such as effect sizes and Bayes factors may provide a more nuanced perspective.

### Effect size and Bayes factor analysis

Effect sizes and Bayes factors improve upon inferences drawn from $p$-values in different ways. The effect size is an estimate of the *magnitude* of an effect, which is stable across sample sizes and allows for a more fine-grained comparison across studies than a binary significance criterion. The Bayes Factor is an alternative hypothesis testing method that quantifies the *evidence* in favor of a hypothesis [@jeffreys1961]. Among other advantages, it does not privilege the null hypothesis and can provide evidence either *for* or *against* [see also @scheibehenne2016, @wagenmakers2016].  While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

These additional measures therefore give another sense in which replications can be successful: if we find roughly the same effect size or Bayes factor despite $p$-values on different sides of the $p = 0.05$ threshold, we should have increased confidence in the effect's reproducibility. To compute a standardized effect size, we converted all one-sample $t$-statistics into a Cohen's $d$ using the formula $d = t/\sqrt{df}$, where $df$ is the degrees of freedom. We first compared effect sizes from the original and replication studies (Figure 2A). Effect sizes were highly correlated between the two sets of studies 
(r = `r signif(cor.test(d$original_ES_d,d$rep_es)$estimate, digits =2)`, p < .0001), but they were generally smaller in the replications 
(`r signif(coef(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)`\% of the original [`r signif(confint(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)` - `r signif(confint(lm(d$rep_es~d$original_ES_d-1))[2], digits = 2)`]). This estimate should be viewed with caution, however, due to its dependence on one very large effect size estimate ($d > 15$) from a single study.

Next, following @etz2016, we compared Bayes Factors between the original and replication. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpretating confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).
